---
title: "Simple Iterations"
author: "Bastola"
date: "`r format(Sys.Date(), ' %B %d %Y')`"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      collapse = TRUE, 
                      comment=NA, 
                      warning = FALSE,
                      message = FALSE,
                      fig.height = 4, fig.width = 6, fig.align='center')

library(tidyverse) 
library(lubridate)
library(stringr)
library(forcats)

energy <- readr::read_csv("https://raw.githubusercontent.com/deepbas/statdatasets/main/energy.csv",
                    col_type = cols(
                     .default = col_double(), 
                      Timestamp = col_datetime(format = ""),
                      dayWeek = col_factor(levels=c("Mon","Tues","Wed","Thurs","Fri","Sat","Sun"))
                     ))

```

## Your Turn 1

Load the `palmerpenguins` package.

Write a `for` loop that calculates the mean of the numeric variables in the `penguins` data set and stores the means in a named vector.

**Solution:** Looking at the `penguins` data set, the columns 3, 4, 5, 6, and 8 could be considered numeric (you could debate the year I suppose).

```{r}
library(palmerpenguins)
glimpse(penguins)
```

We'll use the numeric column index as our vector to iterate over:

```{r}
numeric_cols <- _________ # indexing the numeric columns

col_means <- ________   # empty vector with NAs to store means
names(col_means) <- colnames(penguins) # adding col names

# Calculate means for numeric columns

for(i in numeric_cols) {
  col_means[i] <- _________________
}
col_means
```



---------------------------------------------------


## Your turn 2

**1.** Make a data frame of quantiles for `energy` buildings in columns 9-90 (you will need `na.rm = TRUE`)

```{r}
glimpse(energy)
```



```{r}
qdf <- energy %>% select(9:90) %>%
  _______(quantile, probs = seq(.1,.9,.1), _____)
```


**2.**  Add a variable called `stat` to identify the quantile

```{r}
qdf <- energy %>% select(9:90) %>%
  ______(quantile, probs = seq(.1,.9,.1), _______) %>%  
  ______(stat = ____("quantile_", seq(10,90,10))) 
```


**3.**  Reshape the data frame to make variables `stat` (describing the quantile), `building` and `quant` (quantile value)

```{r}
qdf <- energy %>% select(9:90) %>%
  map_dfc(quantile, probs = seq(.1,.9,.1), na.rm = TRUE) %>%  
  mutate(stat = str_c("quantile_", seq(10,90,10))) %>%   
  ________  # use gather or pivot_longer  
```

**4.**  Plot the KWH value for each quantile on the x-axis for the buildings `Sayles-Hill`, `Language_&_Dining_Center`, `Olin_Hall_of_Science`

```{r}
qdf %>% 
  ______(building %in% c("Sayles-Hill" ,"Language_&_Dining_Center", "Olin_Hall_of_Science"))  %>%  
  ggplot(aes(x=q, y=parse_number(stat), color=building)) +
  geom_point() + 
  geom_line(aes(group=building)) + 
  labs(y="Percentile (%)",x="KWH") +
  scale_y_continuous(breaks=seq(10,90,by=10))
```




